---
title: "ion2cbioportal"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ion2cbioportal}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

ion2cbioportal is a R package for transformation of the ion torrent files to the cbioportal data files. The ion torrent files include a vcf and a tsv files. Both files should be filtered file that means both files includes only filtered variants not all variants. Both files are tied in one compressed files per one sample test. The cbioportal data file has maf and cnv table files (https://github.com/cBioPortal/cbioportal/blob/master/docs/File-Formats.md).

#### Steps  

##### 1. Extract vcf and tsv from a compressed files  
```{r extract_ion2cbio}
extract_ion2cbio("/home/molpath/cbioportal_data/zip", "/home/molpath/cbioportal_data/output")
```

Two subdirectories (vcf and tsv) will be created and include files.  

##### 2. Transform vcf to maf  

Go to terminal and run shell script `/home/molpath/Desktop/temp_2019_1.sh`. Then maf directory will be created in `/home/molpath/cbioportal_data/maf/`.  

Below does not work. NOTE ONLY.

```{r vcf2maf}
vcf_list <- list.files("/home/molpath/cbioportal_data/output/vcf/", full.names = TRUE)
m_id <- stringr::str_extract(vcf_list, 'M[0-9]{2}-[0-9]{1,8}')
program_vcf2maf <- "/home/molpath/mskcc-vcf2maf-decbf60/vcf2maf.pl"
arg1 <- paste("--input-vcf", vcf_list)
arg2 <- paste0("--output-maf /home/molpath/cbioportal_data/maf/", m_id, ".maf")
arg3 <- paste("--tumor-id", m_id)

vcf2maf_script <- paste(program_vcf2maf, arg1, arg2, arg3)
old_path <- Sys.getenv("PATH")
Sys.setenv(PATH = paste(old_path, "/home/molpath/samtools-1.8/", sep = ":"))
system(vcf2maf_script[1]) 
```


##### 3. Merge maf files  

Because the cbioportal maf file should include all data in one maf file, maf files of multiple samples must be merged. The `all_maf(/path/to/maf/need/to/merge)` generates merged maf file (`output/directory/all_maftools.maf`). This is the final cbioportal maf data file.  

```{r merge maf}
all_maf("./maf", "/output/directory")
```


##### 4. Transform tsv to cnv data file  

The tsv file contains amplification data of ion torrent. The `merge_cnv` function makes cnv table for cbioportal cnv log2 data file. The `discreize_cnv` transforms cnv log2 data to the discrete cnv data (0,2).

```{r merge cnv}
table_cnv <- merge_cnv(f_list)
discrete_cnv <- discretize_cnv(table_cnv)
write.table(table_cnv, './example_data/data_log2_CNA.txt', append = FALSE, quote = FALSE, sep = "\t", na = "NA", dec = ".", col.names = TRUE, row.names = FALSE)
write.table(discrete_table, './example_data/data_CNA.txt', append = FALSE, quote = FALSE, sep = "\t", na = "NA", dec = ".", col.names = TRUE, row.names = FALSE)
```

##### 5. Make case list files  

`make_case_list` generates `Case list files` in `study-dir/case_lists` directory.  

```{r}
make_case_list("~/cbioportal_data/maf")
```

##### 6. Append new data to old data  

##### 6.1. Clinical data file  

```{r}
append_patient(new_patient_file, old_patient_file)
append_sample(new_sample_file, old_sampe_file)
```

Checklist of the patient and sample files  

- Is each data row ended without white space
- Is the last line ended with /n
- No additional empty data after the last line

###### 6.2. MAF file  

```{r}
append_all_maf(new_maf_file, old_maf_file)
```

###### 6.3 CNV discrete file  

```{r}
append_discrete_cnv(new_discrete_cnv_table, old_discrete_cnv_file)
```

###### 6.4 Case list files  

```{r}
append_list(new_list_directory, old_list_directory)
```


The appended data files will override the original data files. It is safe to copy the last data set and override the original data.  

##### 7. Import cbioportal data  

If cbioportal is deployed with thehyve docker image, follow the instructions. https://github.com/thehyve/cbioportal-docker/blob/master/docs/example_commands.md `-o` option means overriding warnings.  


```{}
cd ~/cbioportal-docker-compose

docker-compose run \
    -v /home/humintec/cbioportal_report:/report \
    cbioportal \
    metaImport.py -u http://cbioportal:8080 -s /study/tnbcCMC -o --html=/report/report.html

docker-compose restart cbioportal
docker restart cbioportal-container
```

##### 7. Lessons  

1. The thehyve is better than the cbioportal documentation for deploying cbioportal using docker. The cbioportal documentation makes problem in migration of database to latest version.  

2. If migration of the database to latest version fails, the undefined cancer type error will follow.  

3. vcf2maf https://github.com/mskcc/vcf2mafInstall 
Download and unpack VEP's offline cache for GRCh37, GRCh38, and GRCm38: This download takes very long time (3 days).  

4. Vignette will be installed with `devtools::build_vignettes()` -> `devtools::check()` -> `devtools::build()` -> install.package from source `install.packages("~/ion2cbioportal_0.1.0.tar.gz", repos = NULL, type = "source")`

5. Example in roxygen comments will be checked whether it can be run during `devtools::check()`. If it fails executing, error will follow. When the example command is not executable in the given environment, use `\dontrun{}`.  

6. Some packages such as dplyr, tibble, utils need to import at function documentation arguments ```#' @import dplyr```  